\textbf{SPECFEM3D Cartesian}\\ \textbf{User Manual}

### 2014-06-19 doc/USER_MANUAL/manual_SPECFEM3D_Cartesian.tex

The basic steps in creating a load-balanced, partitioned mesh with
CUBIT are:
\begin{description}
\item [{1.}] setting up a hexahedral mesh with CUBIT,
\item [{2.}] exporting the CUBIT mesh into a SPECFEM3D Cartesian file format
and
\item [{3.}] partitioning the SPECFEM3D Cartesian mesh files for a chosen
number of cores.
\end{description}
Examples are provided in the SPECFEM3D Cartesian package in the subdirectory
\texttt{examples/}. We strongly encourage you to contribute your own
example to this package by contacting the CIG Computational Seismology
Mailing List \urlwithparentheses{cig-seismo@geodynamics.org}.

@@

For the installation and handling of the CUBIT meshing tool suite,
please refer to the CUBIT user manual and documentation. In order
to give you a basic understanding of how to use CUBIT for our purposes,
examples are provided in the SPECFEM3D Cartesian package in the subdirectory
\texttt{examples/}:
\begin{description}
\item [{\texttt{homogeneous\_halfspace}}] Creates a single block model
and assigns elastic material parameters.
\item [{\texttt{layered\_halfspace}}] Combines two different, elastic material
volumes and creates a refinement layer between the two. This example
can be compared for validation against the solutions provided in subdirectory
~\\
 \texttt{VALIDATION\_3D\_SEM\_SIMPLER\_LAYER\_SOURCE\_DEPTH/}.
\item [{\texttt{waterlayered\_halfspace}}] Combines an acoustic and elastic
material volume as in a schematic marine survey example.
\item [{\texttt{tomographic\_model}}] Creates a single block model whose
material properties will have to be read in from a tomographic model
file during the databases creation by \texttt{xgenerate\_databases}.
\end{description}
\begin{figure}[htbp]
\noindent \begin{centering}
\includegraphics[width=0.45\textwidth]{figures/example-homogeneous}
\includegraphics[width=0.45\textwidth]{figures/example-2layers} \\
 \includegraphics[width=0.45\textwidth]{figures/example-water} \includegraphics[width=0.45\textwidth]{figures/example-tomo}
\par\end{centering}

\caption{Screenshots of the CUBIT examples provided in subdirectory \texttt{examples/}:
homogeneous halfspace (top-left), layered halfspace (top-right), water
layered halfspace (bottom-left) and tomographic model (bottom-right).}

@@

IMPORTANT note: in order to correctly set up GEOCUBIT and run the examples, please read the file called "examples/README";
in particular, please make sure you correctly set up the Python paths as indicated in that file.

@@

\begin{lyxcode}
NEX\_XI\_BEGIN~NEX\_XI\_END~NEX\_ETA\_BEGIN~NEX\_ETA\_END~NZ\_BEGIN~NZ\_END~material\_ID
\end{lyxcode}
\end{description}
The \texttt{INTERFACES\_FILE} parameter of \texttt{Mesh\_Par\_File}
defines the file which contains the settings of the topography grid
and of the interfaces grids. Topography is defined as a set of elevation
values on a regular 2D grid. It is also possible to define interfaces
between the layers of the model in the same way. The file needs to
define several parameters:
\begin{itemize}
\item The number of interfaces, including the topography. This needs to
be set at the first line. Then, from the bottom to the top of the
model, you need to define the grids with:
\item \texttt{SUPPRESS\_UTM\_PROJECTION} flag as described previously,
\item number of points along $x$ and $y$ direction (NXI and NETA),
\item minimal $x$ and $y$ coordinates (LONG\_MIN and LAT\_MIN),
\item spacing between points along $x$ and $y$ (SPACING\_XI and SPACING\_ETA)
and
\item the name of the file which contains the elevation values (in $y$.$x$
increasing order).
\end{itemize}
At the end of this file, you simply need to set the number of spectral
elements in the vertical direction for each layer. We provide a few
models in the {\texttt{examples/}} directory. \\

@@

The CUBIT scripts ({*}.jou and {*}.py) in the directory $\mathtt{examples}$
generate more complicated meshes. The {*}.py files are Python scripts
that execute CUBIT commands and use the CUBIT-python interface for
SPECFEM3D (see next section). The Python language allows to define
and manipulate variables to parameterize the mesh. Alternatively,
the Python script can call a CUBIT journal file ({*}.jou), which looks
like the example above. Variables can be defined and manipulated there
using the $\mathtt{APREPRO}$ language built in CUBIT.\\

@@

\end{itemize}
Read the documents in the directory $\mathtt{examples/*/description}$.
They contain a description of the example and additional instructions
to run it. Visualize the results with the Matlab scripts in the directory
$\mathtt{examples/*/post}$

@@

\begin{lyxlist}{00.00.0000}
\item [{$\mathtt{examples/tpv16/description/TPV16\_17\_Description\_v03.pdf}$}]~
\end{lyxlist}

@@

The directories $\mathtt{examples/*/post}$ contain additional Matlab
scripts to generate figures specific to each example.

@@

\texttt{examples/noise\_tomography/NOISE\_TOMOGRAPHY.m} (main program)\\
 \texttt{examples/noise\_tomography/PetersonNoiseModel.m}

@@

In order to illustrate noise simulations in an easy way, one example
is provided in \texttt{examples/noise\_tomography/}. See \texttt{examples/noise\_tomography/README}
for explanations. \\

@@

If your workstation is suitable, you can run the example in \texttt{examples/noise\_tomography/}
using:\\

@@

\item [{after executing \texttt{xmeshfem3D} I've got elements with skewness of 81\% percent, what does this mean:}] Look
at the skewness table printed in the \texttt{output\_mesher.txt} file
after executing \texttt{xmeshfem3D} for the example given in \texttt{examples/meshfem3D\_examples/simple\_model/}:

@@

To give you an idea why some of the elements are distorted, see the
following figure \ref{fig:mesh.vp} of the mesh you obtain in the
example \texttt{examples/meshfem3D\_examples/simple\_model/}.
\begin{figure}[htbp]
\noindent \begin{centering}
\includegraphics[width=0.9\textwidth]{figures/mesh_vp}
\par\end{centering}

\caption{Paraview visualization using the mesh vtk-files for the example given
in \texttt{examples/meshfem3D\_examples/simple\_model/}.}

>>>>

The basic steps in creating a load-balanced, partitioned mesh with
CUBIT are:
\begin{description}
\item [{1.}] setting up a hexahedral mesh with CUBIT,
\item [{2.}] exporting the CUBIT mesh into a SPECFEM3D Cartesian file format
and
\item [{3.}] partitioning the SPECFEM3D Cartesian mesh files for a chosen
number of cores.
\end{description}
Examples are provided in the SPECFEM3D Cartesian package in the subdirectory
\texttt{EXAMPLES/}. We strongly encourage you to contribute your own
example to this package by contacting the CIG Computational Seismology
Mailing List \urlwithparentheses{cig-seismo@geodynamics.org}.

@@

For the installation and handling of the CUBIT meshing tool suite,
please refer to the CUBIT user manual and documentation. In order
to give you a basic understanding of how to use CUBIT for our purposes,
examples are provided in the SPECFEM3D Cartesian package in the subdirectory
\texttt{EXAMPLES/}:
\begin{description}
\item [{\texttt{homogeneous\_halfspace}}] Creates a single block model
and assigns elastic material parameters.
\item [{\texttt{layered\_halfspace}}] Combines two different, elastic material
volumes and creates a refinement layer between the two. This example
can be compared for validation against the solutions provided in subdirectory
~\\
 \texttt{VALIDATION\_3D\_SEM\_SIMPLER\_LAYER\_SOURCE\_DEPTH/}.
\item [{\texttt{waterlayered\_halfspace}}] Combines an acoustic and elastic
material volume as in a schematic marine survey example.
\item [{\texttt{tomographic\_model}}] Creates a single block model whose
material properties will have to be read in from a tomographic model
file during the databases creation by \texttt{xgenerate\_databases}.
\end{description}
\begin{figure}[htbp]
\noindent \begin{centering}
\includegraphics[width=0.45\textwidth]{figures/example-homogeneous}
\includegraphics[width=0.45\textwidth]{figures/example-2layers} \\
 \includegraphics[width=0.45\textwidth]{figures/example-water} \includegraphics[width=0.45\textwidth]{figures/example-tomo}
\par\end{centering}

\caption{Screenshots of the CUBIT examples provided in subdirectory \texttt{EXAMPLES/}:
homogeneous halfspace (top-left), layered halfspace (top-right), water
layered halfspace (bottom-left) and tomographic model (bottom-right).}

@@

IMPORTANT note: in order to correctly set up GEOCUBIT and run the examples, please read the file called "EXAMPLES/README";
in particular, please make sure you correctly set up the Python paths as indicated in that file.

@@

\begin{lyxcode}
NEX\_XI\_BEGIN~NEX\_XI\_END~NEX\_ETA\_BEGIN~NEX\_ETA\_END~NZ\_BEGIN~NZ\_END~material\_ID
\end{lyxcode}
\end{description}
The \texttt{INTERFACES\_FILE} parameter of \texttt{Mesh\_Par\_File}
defines the file which contains the settings of the topography grid
and of the interfaces grids. Topography is defined as a set of elevation
values on a regular 2D grid. It is also possible to define interfaces
between the layers of the model in the same way. The file needs to
define several parameters:
\begin{itemize}
\item The number of interfaces, including the topography. This needs to
be set at the first line. Then, from the bottom to the top of the
model, you need to define the grids with:
\item \texttt{SUPPRESS\_UTM\_PROJECTION} flag as described previously,
\item number of points along $x$ and $y$ direction (NXI and NETA),
\item minimal $x$ and $y$ coordinates (LONG\_MIN and LAT\_MIN),
\item spacing between points along $x$ and $y$ (SPACING\_XI and SPACING\_ETA)
and
\item the name of the file which contains the elevation values (in $y$.$x$
increasing order).
\end{itemize}
At the end of this file, you simply need to set the number of spectral
elements in the vertical direction for each layer. We provide a few
models in the {\texttt{EXAMPLES/}} directory. \\

@@

The CUBIT scripts ({*}.jou and {*}.py) in the directory $\mathtt{EXAMPLES}$
generate more complicated meshes. The {*}.py files are Python scripts
that execute CUBIT commands and use the CUBIT-python interface for
SPECFEM3D (see next section). The Python language allows to define
and manipulate variables to parameterize the mesh. Alternatively,
the Python script can call a CUBIT journal file ({*}.jou), which looks
like the example above. Variables can be defined and manipulated there
using the $\mathtt{APREPRO}$ language built in CUBIT.\\

@@

\end{itemize}
Read the documents in the directory $\mathtt{EXAMPLES/*/description}$.
They contain a description of the example and additional instructions
to run it. Visualize the results with the Matlab scripts in the directory
$\mathtt{EXAMPLES/*/post}$

@@

\begin{lyxlist}{00.00.0000}
\item [{$\mathtt{EXAMPLES/tpv16/description/TPV16\_17\_Description\_v03.pdf}$}]~
\end{lyxlist}

@@

The directories $\mathtt{EXAMPLES/*/post}$ contain additional Matlab
scripts to generate figures specific to each example.

@@

\texttt{EXAMPLES/noise\_tomography/NOISE\_TOMOGRAPHY.m} (main program)\\
 \texttt{EXAMPLES/noise\_tomography/PetersonNoiseModel.m}

@@

In order to illustrate noise simulations in an easy way, one example
is provided in \texttt{EXAMPLES/noise\_tomography/}. See \texttt{EXAMPLES/noise\_tomography/README}
for explanations. \\

@@

If your workstation is suitable, you can run the example in \texttt{EXAMPLES/noise\_tomography/}
using:\\

@@

\item [{after executing \texttt{xmeshfem3D} I've got elements with skewness of 81\% percent, what does this mean:}] Look
at the skewness table printed in the \texttt{output\_mesher.txt} file
after executing \texttt{xmeshfem3D} for the example given in \texttt{EXAMPLES/meshfem3D\_examples/simple\_model/}:

@@

To give you an idea why some of the elements are distorted, see the
following figure \ref{fig:mesh.vp} of the mesh you obtain in the
example \texttt{EXAMPLES/meshfem3D\_examples/simple\_model/}.
\begin{figure}[htbp]
\noindent \begin{centering}
\includegraphics[width=0.9\textwidth]{figures/mesh_vp}
\par\end{centering}

\caption{Paraview visualization using the mesh vtk-files for the example given
in \texttt{EXAMPLES/meshfem3D\_examples/simple\_model/}.}
